\myTitle

### 2024-12-10 cover-letter/responses-to-editor-and-referees.Rmd

---
output: pdf_document
---

Benjamin Sibbett PhD  
Executive Editor, Molecular Ecology Resources  
4 December 2024

Dear Dr Sibbett,

Thank you for the opportunity to revise our work for consideration of acceptance
in _Molecular Ecology Resources_.  We appreciate the efforts and helpful
questions and suggestions from the associate editor and the two referees.  We have
addressed all queries in the decision letter and are pleased with how doing so
has improved the manuscript.

On the following pages of thie PDF document, please find:

- a PDF version of our responses to the decision letter (these are also pasted in 
into the text box on manuscript central)
- a latexdiff PDF file showing changes we made to between the original submission and our
new revised version of the manuscript.

Thank you again for considering our work.  As always, the MER editorial process has
been a model of efficiency and clarity.  Thank you for all your efforts and leadership.

Best regards,

\vspace*{4em}

Eric C. Anderson, PhD

\newpage

# Responses to Editor and Referees

@@

L120-123: Is it not possible to simply filter the loci set based on linkage disequilibrium? Do you have evidence that this will not alleviate the problem?

_No, it is not possible. Genomic material is inherited in chunks that are typically orders of magnitude longer than the typical range over which LD decays.  Thus, pairs of markers that are on the same chromosome, are much more likely to be co-inherited than markers that are not on the same chromosome, even if the two markers are far enough apart on the same chromosome that they are not in LD. We have noted this in the discussion for any other readers that may wonder whether LD pruning will solve the problem of physical linkage._

L196-198: How would the requirement of having no inbreeding loops in the GSP affect the applicability of these simulations to real-life scenarios? Within most wild populations, especially threatened species, inbreeding is a regular occurrence.

_This is a reasonable question and we address it now in the Discussion. In short, gscramble works fine for inbred populations.  Partly this is because if the founders are related (as they would be in a tiny population) the offspring produced by the GSP will be inbred despite the absence of inbreeding loops in the GSP._

@@

L227-229: Will recombination rates be something the user will have to input?

_Yes, the user has to put those in, though we already note in the discussion that even if an accurate recombination map is not known, it is still possible to use one simply with average rates across each chromosome._

L363-365: My understanding was that gscramble accounted for the bias associated with linked loci without the need to filter, so why would an empirical dataset need to be filtered for linked loci? What benefit then does gscramble present over other methods in this regard?

_As noted above, LD and physical linkage are quite different.  Where we now address this issuein the discussion, we also note that for analyses of simulated data produced with gscramble, i.e., the use of ADMIXTURE, it can be advantageous to conduct LD-pruning._

@@

3. Consolidate Supplemental Information into as few files as possible.

    * Use the attached Molecular Ecology Resources Supplemental Information Template if possible (not required)
    * If the file is quite long, create a Table of Contents for the first page
    * Make sure the title of the paper and the names of all authors appear on the Supplemental document.
    * Your Supplemental Information will not be copy-edited.  Do not leave in track-changes and other editing marks.  The document will be posted "as is.”
    * Save as PDF if possible.

_The Supplemental Information has been saved as a PDF_

4. Provide your manuscript in an editable format (not PDF!)
5. Provide tables in an editable format (not images)
6. Make sure all table and figure captions are present and editable

### 2024-12-10 abstract.tex
Revised and resubmitted version

While a best practice for evaluating the behavior of genetic clustering algorithms 
on empirical data is to conduct parallel analyses on simulated data, these types 
of simulation techniques often involve sampling genetic data with replacement. 
In this paper we demonstrate that sampling with replacement, especially with large 
marker sets, inflates the perceived statistical power to correctly assign individuals
(or the alleles that they carry)
back to source populations---a phenomenon we refer to as resampling-induced, 
spurious power inflation (RISPI). To address this issue, we present \gscramble{}, a 
simulation approach in R for creating biologically informed individual genotypes from 
empirical data that: 1) samples alleles from populations \textit{without} replacement, 
2) segregates alleles based on species-specific recombination rates. This framework
makes it possible to simulate admixed individuals in a way that respects the physical
linkage between markers on the same chromosome and which does not suffer
from RISPI.  This is achieved in \gscramble{} by allowing 
users to specify pedigrees of varying complexity 
in order to simulate admixed genotypes, segregating and tracking haplotype blocks 
from different source populations through those pedigrees, and then sampling---using
a variety of permutation schemes---alleles from empirical data into those haplotype blocks.
We demonstrate the functionality of 
\gscramble{} with both simulated and empirical data sets and highlight additional uses of 
the package that users may find valuable.

>>>>

While a best practice for evaluating the behavior of genetic clustering algorithms 
on empirical data is to conduct parallel analyses on simulated data, these types 
of simulation techniques often involve sampling genetic data with replacement. 
In this paper we demonstrate that sampling with replacement, especially with large 
marker sets, inflates the perceived statistical power to correctly assign individuals
(or the alleles that they carry)
back to source populations---a phenomenon we refer to as resampling-induced, 
spurious power inflation (RISPI). To address this issue, we present \gscramble{}, a 
simulation approach in R for creating biologically informed individual genotypes from 
empirical data that: 1) samples alleles from populations \textit{without} replacement and 
2) segregates alleles based on species-specific recombination rates. This framework
makes it possible to simulate admixed individuals in a way that respects the physical
linkage between markers on the same chromosome and which does not suffer
from RISPI.  This is achieved in \gscramble{} by allowing 
users to specify pedigrees of varying complexity 
in order to simulate admixed genotypes, segregating and tracking haplotype blocks 
from different source populations through those pedigrees, and then sampling---using
a variety of permutation schemes---alleles from empirical data into those haplotype blocks.
We demonstrate the functionality of 
\gscramble{} with both simulated and empirical data sets and highlight additional uses of 
the package that users may find valuable.

### 2024-12-10 body-text-of-paper.tex
Revised and resubmitted version

%%%%%%%
\newcommand{\permcap}{\footnotesize An illustration of how \gscramble{} permutes data under
different options.  {\bf (a)}~A picture
representing an original data set of $N=12$ diploids sampled from three populations and typed at $L=15$ markers.
Different colors represent different sampled individuals. Individuals 1--5 are shades of blue and green, indicating that they
are from a single population.  Individuals 6--9 are from another population, (shades of red and orange), while 10--12
are from a third population and are colored shades of violet. Thick vertical black lines separate the different
populations.  Markers are shown in rows. There are 6 markers on the first chromosome, 5 on the second, and 4 on the
third.  Each chromosome is delimited by a thick, horizontal black line. The ``a'' and the ``b'' in the colored cells denote the first and second gene copies within an individual at each marker.  It is worth noting that this representation shows a $2L \times N$ matrix,
which is the transpose of the $N\times 2L$ matrix used to supply genetic marker data to \gscramble{}. {\bf (b)}~Examples of how the {\tt\footnotesize preserve\_individuals} and {\tt\footnotesize preserve\_haplotypes} options of the {\tt\footnotesize segments2markers()}
function affect the permutation of
the genetic data of the original data set.  Panel headers show the option values while the colors and ``a'' and ``b'' values
in the colored cells show the original position of each gene copy in the original sample.  Note that permutation is always
done within the populations.  When {\tt\footnotesize preserve\_individuals = TRUE},  then just the
individuals in their entirety are permuted.  When {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, each pair
of a single chromosome within an individual remains together when permuted. When {\tt\footnotesize preserve\_individuals = FALSE}, gene copies are freely permuted between individuals.  The option {\tt\footnotesize preserve\_haplotypes = TRUE} causes each haplotype on a chromosome to be preserved
while permuting. This option should only be used when the original data set is phased.
}
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/permutations-plot.png}
\end{center}
\caption[\permcap]{\sometimes{\permcap}}
\label{fig:perms}
\end{figure*}
%%%%%%%

@@

We performed 500 replicate simulations with \gscramble{} using either 30 or 65 markers.  Each
simulated data set was then analyzed using {\sc NewHybrids} \citep{anderson2002model}. Individuals not used as founders were given the ``s'' and ``z'' flags, indicating that they were considered training samples.  The remaining 46 individuals were treated as possible hybrids and their genetic data used to calculate the posterior probability that they belonged to one of 10 hybrid categories as shown in Table~\ref{tab:newhybcats}, using 25,000 MCMC sweeps and discarding the first 5,000 as burn-in.
%%%%%%%%%%%%
\begin{table}
\caption{The 10 hybrid categories considered for inference with {\sc NewHybrids} and the
expected frequency of their genotypes having 2 ($f(C,C)$), 1 ($f(C,S)$) or 0 ($f(S,S)$) copies of ancestry from cutthroat trout.}
\label{tab:newhybcats}
\begin{center}
{\small
\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrrr}
\hline\hline
Category	& $f(C,C)$ &	$f(C,S)$ & 	$f(S,S)$ \\ \hline
\input{inputs/newhybs-cats.tsv}
\end{tabular*}
}
\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
\end{center}
\end{table}
%%%%%%%%%%%%

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/newhybs-texed-crop.pdf}
\end{center}
\caption[\nhbpcap]{\sometimes{\nhbpcap}}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences, and this analysis
supports the choice of \citet{rizza2023limited} to limit markers to those

@@

Here, we present \gscramble{}, an R package implementing a simulation
approach for creating biologically informed individual 
genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
1) alleles are sampled from populations without replacement and 
2) alleles are segregated based on species-specific recombination rates. 
Our results show that both 
of these features are especially important for accurately reflecting empirical data from large genetic 
marker sets (thousands of markers), 
which are common from both custom SNP-chip genotyping and next generation sequencing methods. Sampling alleles without 
replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
essential for avoiding RISPI---an issue that is exacerbated as the number of 
genetic markers increases. Additionally, we demonstrate the importance of accounting for 
the physical linkage of markers when large marker sets are used, and that treating physically linked loci as 
independent erroneously reduces the variation in the admixture 
fractions of simulated individuals.  In connection with this, we also stress that simply
pruning markers to eliminate or reduce linkage disequilibrium (LD)
will not solve the problem posed by physical linkage.
Genomic material is inherited in chunks that are typically orders of magnitude longer than
the typical range over which LD decays.  Thus, pairs of markers that are
on the same chromosome, are much more likely to be co-inherited than markers that
are not on the same chromosome, even if the two markers are far enough apart on the
same chromosome that they are not in LD.  That said, for some downstream analyses (after
running \gscramble{}),  LD pruning can be advantageous; for example, ADMIXTURE will run faster on a smaller, pruned data set.

@@

As noted in the methods, the GSP used to simulate with \gscramble{} must not contain
inbreeding loops, since the occurrence of such loops means genes would be
sampled with replacement from the founders, potentially inducing RISPI and incorrectly
inflating the
perceived power for admixture fraction estimation.  One might reasonably ask whether this
limits the application of \gscramble{} in small populations of conservation concern where 
inbreeding occurs; we submit that it does not.  If genetic structure between two
locales is driven by family structure (as it often may be in very small and inbred populations), that
family structure will be reflected in samples taken from each locale---the samples within a locale 
may be close relatives of one another.  If related individuals are used as founders in a GSP, then
the offspring they produce will essentially be inbred, even though the GSP itself contains no
inbreeding loops, because the founders are related.  The inclusion of related founders in this way, then, also appropriately reflects the genetic drift that is causing allele frequency differences between the locales.

>>>>

%%%%%%%
\newcommand{\permcap}{\footnotesize An illustration of how \gscramble{} permutes data under
different options.  {\bf (a)}~A picture
representing an original data set of $N=12$ diploids sampled from three populations and typed at $L=15$ markers.
Different colors represent different sampled individuals. Individuals 1--5 are shades of blue and green, indicating that they
are from a single population.  Individuals 6--9 are from another population, (shades of red and orange), while 10--12
are from a third population and are colored shades of violet. Thick vertical black lines separate the different
populations.  Markers are shown in rows. There are 6 markers on the first chromosome, 5 on the second, and 4 on the
third.  Each chromosome is delimited by a thick, horizontal black line. The ``a'' and the ``b'' in the colored cells denote the first and second gene copies within an individual at each marker.  It is worth noting that this representation shows a $2L \times N$ matrix,
which is the transpose of the $N\times 2L$ matrix used to supply genetic marker data to \gscramble{}. {\bf (b)}~Examples of how the {\tt\footnotesize preserve\_individuals} and {\tt\footnotesize preserve\_haplotypes} options of the {\tt\footnotesize segments2markers()}
function affect the permutation of
the genetic data of the original data set.  Panel headers show the option values while the colors and ``a'' and ``b'' values
in the colored cells show the original position of each gene copy in the original sample.  Note that permutation is always
done within the populations.  When {\tt\footnotesize preserve\_individuals = TRUE},  then just the
individuals in their entirety are permuted.  When {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, each pair
of a single chromosome within an individual remains together when permuted. When {\tt\footnotesize preserve\_individuals = FALSE}, gene copies are freely permuted between individuals.  The option {\tt\footnotesize preserve\_haplotypes = TRUE} causes each haplotype on a chromosome to be preserved
while permuting. This option should only be used when the original data set is phased.
}
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{figures/permutations-plot.pdf}
\end{center}
\caption[\permcap]{\sometimes{\permcap}}
\label{fig:perms}
\end{figure*}
%%%%%%%

@@

We performed 500 replicate simulations with \gscramble{} using either 30 or 65 markers.  Each
simulated data set was then analyzed using {\sc NewHybrids} \citep{anderson2002model}. Individuals not used as founders were given the ``s'' and ``z'' flags, indicating that they were considered training samples.  The remaining 46 individuals were treated as possible hybrids and their genetic data used to calculate the posterior probability that they belonged to one of 10 hybrid categories as shown in Table~\ref{tab:newhybcats}, using 25,000 MCMC sweeps and discarding the first 5,000 as burn-in.
%%%%%%%%%%%%
\begin{table}
\caption{The 10 hybrid categories considered for inference with {\sc NewHybrids} and the
expected frequency of their genotypes having 2 ($f(C,C)$), 1 ($f(C,S)$) or 0 ($f(S,S)$) copies of ancestry from cutthroat trout.}
\label{tab:newhybcats}
\begin{center}
{\small
\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrrr}
\hline\hline
Category	& $f(C,C)$ &	$f(C,S)$ & 	$f(S,S)$ \\ \hline
\input{inputs/newhybs-cats.tsv}
\end{tabular*}
}
\vspace*{-0.3ex}\hrule\vspace*{0.3ex}\hrule
\end{center}
\end{table}
%%%%%%%%%%%%

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
\begin{center}
\includegraphics[width=0.48\textwidth]{figures/newhybs-texed-crop.pdf}
\end{center}
\caption[\nhbpcap]{\sometimes{\nhbpcap}}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more posteriors $>0.90$ for an {\em incorrect} hybrid category
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences, and this analysis
supports the choice of \citet{rizza2023limited} to limit markers to those on different chromosome arms during their {\sc NewHybrids} analysis.

@@

Here, we present \gscramble{}, an R package implementing a simulation
approach for creating biologically informed individual 
genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
1) alleles are sampled from populations without replacement and 
2) alleles are segregated based on species-specific recombination rates. 
Our results show that both 
of these features are especially important for accurately reflecting empirical data from large genetic 
marker sets (thousands of markers), 
which are common from both custom SNP-chip genotyping and next generation sequencing methods. Sampling alleles without 
replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
essential for avoiding RISPI---an issue that is exacerbated as the number of 
genetic markers increases. Additionally, we demonstrate the importance of accounting for 
the physical linkage of markers when large marker sets are used, and that treating physically linked loci as 
independent erroneously reduces the variation in the admixture 
fractions of simulated individuals.  In connection with this, we also stress that simply
pruning markers to eliminate or reduce linkage disequilibrium (LD)
will not solve the problem posed by physical linkage.
Genomic material is inherited in chunks that are typically orders of magnitude longer than
the typical range over which LD decays.  Thus, pairs of markers that are
on the same chromosome, are much more likely to be co-inherited than markers that
are not on the same chromosome, even if the two markers are far enough apart on the
same chromosome that they are not in LD.  That said, for some downstream analyses (after
running \gscramble{}),  LD pruning can be advantageous; for example, ADMIXTURE assumes independent loci and will run faster on a smaller, pruned data set.

@@

As noted in the methods, the GSP used to conduct simulations with \gscramble{} must not contain
inbreeding loops, since the occurrence of such loops means genes would be
sampled with replacement from the founders, potentially inducing RISPI and incorrectly
inflating the
perceived power for admixture fraction estimation.  One might reasonably ask whether this
limits the application of \gscramble{} in small populations of conservation concern where 
inbreeding occurs; we submit that it does not.  If genetic structure between two
locales is driven by family structure (as it often may be in very small and inbred populations), that
family structure will be reflected in samples taken from each locale---the samples within a locale 
may be close relatives of one another.  If related individuals are used as founders in a GSP, then the offspring they produce will essentially be inbred because the founders are related even though the GSP contains to inbreeding loops.  The inclusion of related founders in this way, then, also appropriately reflects the genetic drift that is causing allele frequency differences between the locales.

### 2024-12-10 data-accessibility.tex
Revised and resubmitted version

Data and code for the simulation analyses and steelhead-cutthroat analyses are available from \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} and are archived at \url{https://doi.org/10.5281/zenodo.14270660}.  Data and code for the pig analyses are available from \url{https://github.com/mgdesaix/gscramble-paper-pigs} and are archived
at \url{Get-from-Matt}.

>>>>

Data and code for the simulation analyses and steelhead-cutthroat analyses are available from \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} and are archived at \url{https://doi.org/10.5281/zenodo.14270660}.  Data and code for the pig analyses are available from \url{https://github.com/mgdesaix/gscramble-paper-pigs} and are archived
at \url{https://doi.org/10.5281/zenodo.14271038}.
